True versus apparent shapes of astrophysical bow shocks

### 2016-10-21 papers/Paper1/sec-conic.tex
Incorporate appendix into both papers

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary  
\begin{figure}
\setlength\tabcolsep{0pt}
\begin{tabular}{cc}
\includegraphics[height=0.54\linewidth, trim=30 0 30 0, clip]{ellipse_py} &
\includegraphics[height=0.54\linewidth, trim=60 0 70 0, clip]{hyperbola_py}
\end{tabular}
\label{fig:conics}
\caption{Schematic representation of conic like bowshocks: A) Ellipse and B) Hyperbola The center of the conic is located at a distance $x_0=R_0-a$ from the origin at the $x$ axis. 
The radius of curvature at the axis is $R_c$, and $R_{90}$ is t he radius of the conic in the perpendicular direction to $R_0$.
}
\end{figure}
\begin{figure}
\includegraphics[width=\linewidth]{conic1}
\caption{Example of a family of conic sections, all with the same
  radius of curvature at their apex (marked by white dot) on the
  \(x\)~axis: \(R_c = 2 R_0\). The parameter
  $\theta_c = \pm \tan^{-1} b/a$ describes the family, with negative
  values corresponding to hyperbolae.}
\label{fig:conics-family}
\end{figure}

Due to the similarities between the parametrization of the ellipse and the hyperbola, we can do the following parametrization:
\begin{align}
x &= a\Cos(t)-x_0 \label{eq:par-x}\\ 
y &= b\Sin(t) \label{eq:par-y}
\end{align}
where:
\begin{align}
\Cos(t) &= \left\lbrace \begin{array}{c}
\cos t ~\mathrm{if~ellipse} \\
\cosh t ~\mathrm{if~hyperbola}
\end{array}\right.\\
\Sin(t) &= \left\lbrace \begin{array}{c}
\sin t ~\mathrm{if~ellipse} \\
\sinh t ~\mathrm{if~hyperbola}
\end{array}\right. \\
-\pi < t < \pi \\
R_0 &= a - x_0 \\
a<0 ~ \mathrm{for~hyperbola}
\end{align}
We have set a whole family of curves with a singular parameter denoted as $\tan\theta_c \equiv \frac{b}{a}$. Positive values of this parameter describe closed curves
(i.e. ellipses) and the negative ones describe open curves (i.e. hyperbolas). Particular cases are $\theta_c =\frac{\pi}{4}$, which describes a circle and $\theta_c=0$, which describes
a parabola.
To determine uniquely a given conic in terms of measurable quantities, we use the set of characteristic radii $(R_0,R_c,R_{90})$ in such way the conic parameters $(a,b,\theta_c)$
could be derived from the former. The transformation between the conic parameters and the characteristic radii are given by:
\begin{align}
R_c &= \frac{b^2}{a} \label{eq:Rc-conic}\\ 
R_{90} &= b\left[\pm\left(1 - \frac{(R_0-a)^2}{a^2}\right)\right]^{1/2}\label{eq:r90-conic}
\end{align}
In (\ref{eq:r90-conic}), the positive sign correspond to an elliptic quadric, while the negative sign to an hyperbolic quadric. The inverse transformation of (\ref{eq:Rc-conic})
and (\ref{eq:r90-conic}) is given by:
\begin{align}
\frac{a}{R_0} &= \frac{A}{2A-B^2} \label{eq:a-conic}\\
\frac{|b|}{R_0} &= \frac{A}{\left|2A-B^2\right|^{1/2}}\\
\tan\theta_c &= \frac{2A-B^2}{\left|B^2 - 2A\right|^{1/2}} \label{eq:thc-conic} \\
\mathrm{where:~} B &\equiv \frac{R_{90}}{R_0} \\
A &\equiv \frac{R_c}{R_0}
\end{align}

@@

Once the parametrization is done, we can find the apparent shape of
the shell in the observer's frame, following the procedure explained
in section \ref{sec:projection}.  First of all, the intrinsic 3D shape of the shell is given by:
\begin{align}
x &= a\Cos(t)-x_0 \\ 
y &= b\Sin(t)\cos\phi \\
z &=  b\Sin(t)\sin\phi
\end{align}
The azimutal angle where the line of sight is tangent to the shell is given by equation (\ref{eq:tanphi}) and (\ref{eq:alpha}), then:
\begin{align}
\sin\phi_t &= \frac{b}{a}\tan i\Cot(t) 
\end{align}
where:
\begin{align}
\Cot(t) = \left\lbrace \begin{array}{c}
\cot t ~\mathrm{if~ellipse} \\
-\coth t ~\mathrm{if~hyperbola}
\end{array}\right.
\end{align}

@@

We must show that the apparent shape of the shell in the observer's frame is still the same conic. To do that, we can write equations
(\ref{eq:conic-projected-x}) and (\ref{eq:conic-projected-y}) as follows:
\begin{align}
X' &= a'\Cos t' \label{eq:conic-projected-x-2} \\
Y' &= b'\Sin t' \label{eq:conic-projected-y-2} . 
\end{align}

Comparing, (\ref{eq:conic-projected-x}), (\ref{eq:conic-projected-x-2}), (\ref{eq:conic-projected-y}) and (\ref{eq:conic-projected-y-2}) we find that
\begin{align}
a' &= \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} \\
\Cos(t') &= \frac{a'\Cos(t)}{a\cos i} \\
b' &= b
\end{align} 
Which proves that effectively the projection of a given conic is the same conic. With this knowledge, we can obtain the characteristic radii in
the observer's frame.
\begin{align}
D' &= D\cos i \\
R'_0 &= a' - (a-R_0)\cos i \\
R'_0 &= \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2}  - (a-R_0)\cos i
\end{align}

With  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ we find that:
\begin{align}
\frac{R'_0}{D'}=\frac{R_0}{D}\left(1+A\cot^2\theta_c(f(i;\theta_c)-1) \right)
\end{align}
%Where $A\equiv \frac{R_c}{R_0}$

>>>>

%Instead of the excentricity, we utilize the parameter $\theta_c$ to characterize the different curves, where
%$\tan\theta_c = \frac{b}{a}$,  $b$ and $a$ are the typical parameters of conics. A positive value for $\theta_c$ indicates that the given curve is a closed one, i.e
%an ellipse, while a negative value indicates that is an hyperbola. %Insert figures if neccesary  
\begin{figure}
\setlength\tabcolsep{0pt}
\begin{tabular}{cc}
\includegraphics[height=0.54\linewidth, trim=30 0 30 0, clip]{ellipse_py} &
\includegraphics[height=0.54\linewidth, trim=60 0 70 0, clip]{hyperbola_py}
\end{tabular}
\label{fig:conics}
\caption{Schematic representation of conic like bowshocks: A) Ellipse and B) Hyperbola.
  The center of the conic is located at a distance $x_0=R_0-a$ from the origin at the $x$ axis 
  The radius of curvature at the axis is $R_c$, and $R_{90}$ is t he radius of the conic in the
  perpendicular direction to $R_0$.
}
\end{figure}
\begin{figure}
\includegraphics[width=\linewidth]{conic1}
\caption{Example of a family of conic sections, all with the same
  radius of curvature at their apex (marked by white dot) on the
  \(x\)~axis: \(R_c = 2 R_0\). The parameter
  $\theta_c = \pm \tan^{-1} b/a$ describes the family, with negative
  values corresponding to hyperbolae.}
\label{fig:conics-family}
\end{figure}

Due to the similarities between the parametrization of the ellipse and the hyperbola, we can
do the following parametrization\footnote{The wholee formalism descripted in this section does
  not apply for paraboloids ($\theta_c=0$ case). This case is developed in appendix
  \ref{app:parabola}}:
\begin{align}
x &= a\Cos(t)-x_0 \label{eq:par-x}\\ 
y &= b\Sin(t) \label{eq:par-y}
\end{align}
where:
\begin{align}
\Cos(t) &= \left\lbrace \begin{array}{c}
\cos t ~\mathrm{if~ellipse} \\
\cosh t ~\mathrm{if~hyperbola}
\end{array}\right.\\
\Sin(t) &= \left\lbrace \begin{array}{c}
\sin t ~\mathrm{if~ellipse} \\
\sinh t ~\mathrm{if~hyperbola}
\end{array}\right. \\
-\pi < t < \pi \\
R_0 &= a - x_0 \\
a<0 ~ \mathrm{for~hyperbola}
\end{align}
We have set a whole family of curves with a singular parameter denoted as $\tan\theta_c \equiv \frac{b}{a}$. Positive values of this parameter describe closed curves
(i.e. ellipses) and the negative ones describe open curves (i.e. hyperbolas). Particular cases are $\theta_c =\frac{\pi}{4}$, which describes a circle and $\theta_c=0$, which describes
a parabola.
To determine uniquely a given conic in terms of measurable quantities, we use the set of characteristic radii $(R_0,R_c,R_{90})$ in such way the conic parameters $(a,b,\theta_c)$
could be derived from the former. The transformation between the conic parameters and the characteristic radii are given by:
\begin{align}
R_c &= \frac{b^2}{a} \label{eq:Rc-conic}\\ 
R_{90} &= b\left[\pm\left(1 - \frac{(R_0-a)^2}{a^2}\right)\right]^{1/2}\label{eq:r90-conic}
\end{align}
In (\ref{eq:r90-conic}), the positive sign correspond to an elliptic quadric, while the negative sign to an hyperbolic quadric. The inverse transformation of (\ref{eq:Rc-conic})
and (\ref{eq:r90-conic}) is given by:
\begin{align}
\frac{a}{R_0} &= \frac{A}{2A-B^2} \label{eq:a-conic}\\
\frac{|b|}{R_0} &= \frac{A}{\left|2A-B^2\right|^{1/2}}\\
\tan\theta_c &= \frac{2A-B^2}{\left|B^2 - 2A\right|^{1/2}} \label{eq:thc-conic} \\
\mathrm{where:~} B &\equiv \frac{R_{90}}{R_0} \\
A &\equiv \frac{R_c}{R_0}
\end{align}

@@

Once the parametrization is done, we can find the apparent shape of
the shell in the observer's frame, following the procedure explained
in section \ref{sec:projection}.  First of all, the intrinsic 3D shape of the shell is given by:
\begin{align}
x &= a\Cos(t)-x_0 \\ 
y &= b\Sin(t)\cos\phi \\
z &=  b\Sin(t)\sin\phi
\end{align}
The azimutal angle where the line of sight is tangent to the shell is given by equation (\ref{eq:tanphi}) and (\ref{eq:alpha}), then:
\begin{align}
\sin\phi_t &= \frac{b}{a}\tan i\Cot(t) 
\end{align}
where:
\begin{align}
\Cot(t) = \left\lbrace \begin{array}{c}
-\cot t ~\mathrm{if~ellipse} \\
\coth t ~\mathrm{if~hyperbola}
\end{array}\right.
\end{align}

@@

If the apparent shape of the shell in the observer's frame is still the same conic, then, we
can write equations(\ref{eq:conic-projected-x}) and (\ref{eq:conic-projected-y}) as follows:
\begin{align}
X' = a'\Cos( t') \label{eq:conic-projected-x-2} \\
Y' = b'\Sin (t') \label{eq:conic-projected-y-2} . 
\end{align}

Comparing, (\ref{eq:conic-projected-x}), (\ref{eq:conic-projected-x-2}), (\ref{eq:conic-projected-y}) and (\ref{eq:conic-projected-y-2}) we find that
\begin{align}
a' &= \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2} \\
\Cos(t') &= \frac{a'\Cos(t)}{a\cos i} \\
b' &= b
\end{align}

\begin{align}
D' &= D\cos i \\
R'_0 &= a' - (a-R_0)\cos i \\
R'_0 &= \left(a^2\cos^2 i \pm b^2\sin^2 i\right)^{1/2}  - (a-R_0)\cos i
\end{align}

With  $f(i;\theta_c)\equiv\left(1\pm\tan^2\theta_c\tan^2i\right)^{1/2}$ we find that:
\begin{align}
\frac{R'_0}{D'}=\frac{R_0}{D}\left(1\pm A\cot^2\theta_c(f(i;\theta_c)-1) \right)
\end{align}
%Where $A\equiv \frac{R_c}{R_0}$

### 2016-10-21 papers/Paper1/sec-thin-shell.tex
Incorporate appendix into both papers

\section{Hypersonic thin shell model}
\label{sec:crw-scenario}
This scenario can be described with the formalism of
\citet[][hereafter \CRW{}]{Canto:1996}, who proposes an algebraic
solution for the two winds interaction problem in the thin shell
approximation, in terms of a free parameter
$\beta\equiv\frac{\dot{M}_wv_w}{\dot{M}_{w1}v_{w1}}$, the winds
momentum ratio, which the quantities with subindex 1 corresponds to
the strongest wind, therefore $\beta<=1$. In this work the weakest wind is modeled as an anisotropic hemispherical radial wind
with the following density distribution:

@@

\begin{align}
R_{90} = D\tan\theta_{1,90} \\
\theta_{1,90}\cot\theta_{1,90} = 1-\frac{2\beta}{k+2} \label{eq:th190}
\end{align}
Where $\theta_{1,90}\equiv \theta_1(\frac{\pi}{2})$. Combining both equations and  introducing the parameter 
$\xi\equiv \frac{2}{k+2}$ we have:

@@

>>>>

\section{Application to the hypersonic thin shell solution}
\label{sec:crw-scenario}
As an example application of the ideas of this paper, we consider a
generalization of the thin-shell solution presented in
\citet[][hereafter \CRW{}]{Canto:1996}, for the interaction of two
hypersonic, constant velocity stellar winds.  It is assumed that the
two shocks are highly radiative and that the post-shock flows are
perfectly mixed to form a single shell of negligible thickness.  In
this approximation, the shape of the shell was found algebraically by
\CRW{} from considering conservation of linear and angular momentum,
following an approach first outlined in \citet{Wilkin:1996a}.  The
shape of the bow shock shell is uniquely determined by the parameter
\(\beta\), defined as the momentum-loss ratio of the two winds:
\begin{equation}
  \label{eq:beta-definition}
  \beta \equiv \frac{\dot{M}_w V_w}{\dot{M}_{w1} V_{w1}} . 
\end{equation}
In this definition we use the terminology of \CRW{}, in which
\(\dot{M}_w\) and \(V_w\) are the mass-loss rate and velocity,
respectively, of the wind with source at the origin, while
\(\dot{M}_{w1}\) and \(V_{w1}\) are the mass-loss rate and velocity of
the second wind, located at a distance \(D\) along their \(z\) axis.
Note that \CRW{}'s \(z\) and \(r\) correspond to \(x\) and \(y\) in
sections~\ref{sec:generic-model} to \ref{sec:conic} of the current
paper.  By always placing the weaker of the two winds at the origin,
it is only necessary to consider \(\beta \le 1\).

In this work the weakest wind is modeled as an anisotropic hemispherical radial wind
with the following density distribution:

@@

\begin{align}
R_{90} = D\tan\theta_{1,90} \\
\theta_{1,90}\cot\theta_{1,90} = 1-\frac{2\beta}{k+2} \label{eq:th190}
\end{align}
Where $\theta_{1,90}\equiv \theta_1(\frac{\pi}{2})$. Combining both equations and  introducing the parameter 
<$\xi\equiv \frac{2}{k+2}$ we have:

@@

\subsection{Fits to the tail}
\label{sec:fits-tail}

For the hyperbola ``center''
\begin{multline}
  \label{eq:tail-analytic-x0}
  x_{0,\mathrm{t}} = 0.7 \beta^{-0.55} \biggl[
    C_3 \bigl(\log_{10}\beta\bigr)^3 + C_2 \bigl(\log_{10}\beta\bigr)^2 
  \\ + C_1 \log_{10}\beta + C_0
  \biggr]
\end{multline}
\begin{equation}
  \label{eq:tail-analytic-x0-minus-a}
  (x_{0,\mathrm{t}} - a_{\mathrm{t}}) = D_2 (\log_{10}\beta)^2 + D_1 \log_{10}\beta + D_0
\end{equation}
where
\begin{alignat}{2}
  C_k &= c_{2,k} \xi^2 + c_{1,k} \xi + c_{0,k} &\quad \text{for\ } k &= \{0, 1, 2, 3\} \\
  D_k &= d_{2,k} \xi^2 + d_{1,k} \xi + d_{0,k} &\quad \text{for\ } k &= \{0, 1, 2\} \\
  \label{eq:tail-analytic-coeffs}
\end{alignat}

\newcommand\iso{\ensuremath{^{\mathrm{iso}}}}

\begin{table}
  \caption{Coefficients for hyperbola fit to bowshock tails}
  \label{tab:tail-fit-coeffs}
  \begin{align*}
  C\iso_0 & = {0.123}  & c_{0,0} & = {0.123} & c_{1,0} & = {0.123} & c_{2,0} & = {0.123} \\
  C\iso_1 & = {0.123}  & c_{0,1} & = {0.123} & c_{1,1} & = {0.123} & c_{2,1} & = {0.123} \\
  C\iso_2 & = {0.123}  & c_{0,2} & = {0.123} & c_{1,2} & = {0.123} & c_{2,2} & = {0.123} \\
  C\iso_3 & = {0.123}  & c_{0,3} & = {0.123} & c_{1,3} & = {0.123} & c_{2,3} & = {0.123} \\
  D\iso_0 & = {0.123}  & d_{0,0} & = {0.123} & d_{1,0} & = {0.123} & d_{2,0} & = {0.123} \\
  D\iso_1 & = {0.123}  & d_{0,1} & = {0.123} & d_{1,1} & = {0.123} & d_{2,1} & = {0.123} \\
  D\iso_2 & = {0.123}  & d_{0,2} & = {0.123} & d_{1,2} & = {0.123} & d_{2,2} & = {0.123} \\
  \end{align*}
\end{table}

\begin{figure*}
  \centering
  \includegraphics[width=\linewidth]{figs/conic-head-tail-fit}
  \caption{Double quadric fits to thin shell
    solutions. \textit{Replace these with a version that uses analytic
    fit to tail hyperbola parameters}}
  \label{fig:head-tail}
\end{figure*}
